ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-10-02 pyked.tex
Start more thorough description of Python examples
Work on #40

The \ck{} file that will be used in this example can be found in the
\texttt{tests} directory of the PyKED
repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_rcm.yaml}}
Examining that file, we find the first section specifies the information about
the \ck{} file itself:
%
\begin{yamlbox}
file-author:
  name: Kyle E Niemeyer
  ORCID: 0000-0003-4425-7097
file-version: 0
chemked-version: 0.1.6
\end{yamlbox}
%
Then, we find the information regarding the article in the literature from which
this data was taken. In this case, the dataset comes from the work of
\textcite{Mittal2006a}:
%r
\begin{yamlbox}
reference:
  doi: 10.1002/kin.20180
  authors:
    - name: Gaurav Mittal
    - name: Chih-Jen Sung
      ORCID: 0000-0003-2046-8076
    - name: Richard A Yetter
  journal: International Journal of Chemical Kinetics
  year: 2006
  volume: 38
  pages: 516-529
  detail: Fig. 6, open circle
experiment-type: ignition delay
apparatus:
  kind: rapid compression machine
  institution: Case Western Reserve University
  facility: CWRU RCM
\end{yamlbox}
%
Finally, this file contains just a single datapoint, which describes the experimental
ignition delay, initial mixture composition, initial temperature, initial pressure,
compression time, ignition type, and volume history that specifies
how the volume of the reactor varies with time, for simulating the compression
stroke and post-compression processes.
%
\begin{yamlbox}
datapoints:
- temperature:
    - 297.4 kelvin
  ignition-delay:
    - 1.0 ms
  pressure:
    - 958.0 torr
  composition:
    kind: mole fraction
    species:
      - species-name: H2
        InChI: 1S/H2/h1H
        amount:
          - 0.12500
      - species-name: O2
        InChI: 1S/O2/c1-2
        amount:
          - 0.06250
      - species-name: N2
        InChI: 1S/N2/c1-2
        amount:
          - 0.18125
      - species-name: Ar
        InChI: 1S/Ar
        amount:
          - 0.63125
  ignition-type:
    target: pressure
    type: d/dt max
  compression-time:
    - 38.0 ms
  volume-history:
    time:
      units: s
      column: 0
    volume:
      units: cm3
      column: 1
    values:
      - [0.00E+000, 5.47669375000E+002]
      - [1.00E-003, 5.46608789894E+002]
      - [2.00E-003, 5.43427034574E+002]
      ...
\end{yamlbox}
%
The values for the \yabox|volume-history| are truncated here to save space. One application of the
data stored in this file is to perform a simulation using Cantera~\autocite{Cantera:2.3.0} to
calculate the ignition delay, including the facility-dependent effects represented in the volume
trace. All information required to perform this simulation is present in the \ck{} file, with the
exception of a chemical kinetic model for \ce{H2}\slash {CO} combustion.
%
\begin{pythonbox}
import cantera as ct
from pyked import ChemKED

# Load the ChemKED file and retrieve the first element of the
# datapoints list, which is an instance of the DataPoint class
ck = ChemKED('testfile_rcm.yaml')
dp = ck.datapoints[0]
\end{pythonbox}
%
The initial temperature, pressure, and mixture composition can be read from the
instance of the \pybox|DataPoint| class. \pk{} uses instances of the Pint \pybox|Quantity| class to
store values with units, while Cantera expects a floating-point value in SI
units as input. Therefore, we use the built-in capabilities of Pint to convert
the units from those specified in the \ck{} file to SI units:
%
\begin{pythonbox}
T_initial = dp.temperature.to('K').magnitude
P_initial = dp.pressure.to('Pa').magnitude
X_initial = dp.get_cantera_mole_fraction()

>>>>

The \ck{} file that will be used in this example can be found in the
\texttt{tests} directory of the PyKED
repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_rcm.yaml}}
Examining that file, we find the first section specifies the information about
the \ck{} file itself:
%
\begin{yamlbox}
file-author:
  name: Kyle E Niemeyer
  ORCID: 0000-0003-4425-7097
file-version: 0
chemked-version: 0.1.6
\end{yamlbox}
%
Then, we find the information regarding the article in the literature from which
this data was taken. In this case, the dataset comes from the work of
\textcite{Mittal2006a}:
%r
\begin{yamlbox}
reference:
  doi: 10.1002/kin.20180
  authors:
    - name: Gaurav Mittal
    - name: Chih-Jen Sung
      ORCID: 0000-0003-2046-8076
    - name: Richard A Yetter
  journal: International Journal of Chemical Kinetics
  year: 2006
  volume: 38
  pages: 516-529
  detail: Fig. 6, open circle
experiment-type: ignition delay
apparatus:
  kind: rapid compression machine
  institution: Case Western Reserve University
  facility: CWRU RCM
\end{yamlbox}
%
Finally, this file contains just a single datapoint, which describes the experimental
ignition delay, initial mixture composition, initial temperature, initial pressure,
compression time, ignition type, and volume history that specifies
how the volume of the reactor varies with time, for simulating the compression
stroke and post-compression processes:
%
\begin{yamlbox}
datapoints:
- temperature:
    - 297.4 kelvin
  ignition-delay:
    - 1.0 ms
  pressure:
    - 958.0 torr
  composition:
    kind: mole fraction
    species:
      - species-name: H2
        InChI: 1S/H2/h1H
        amount:
          - 0.12500
      - species-name: O2
        InChI: 1S/O2/c1-2
        amount:
          - 0.06250
      - species-name: N2
        InChI: 1S/N2/c1-2
        amount:
          - 0.18125
      - species-name: Ar
        InChI: 1S/Ar
        amount:
          - 0.63125
  ignition-type:
    target: pressure
    type: d/dt max
  compression-time:
    - 38.0 ms
  volume-history:
    time:
      units: s
      column: 0
    volume:
      units: cm3
      column: 1
    values:
      - [0.00E+000, 5.47669375000E+002]
      - [1.00E-003, 5.46608789894E+002]
      - [2.00E-003, 5.43427034574E+002]
      ...
\end{yamlbox}
%
The values for the \yabox|volume-history| are truncated here to save space. One application of the
data stored in this file is to perform a simulation using Cantera~\autocite{Cantera:2.3.0} to
calculate the ignition delay, including the facility-dependent effects represented in the volume
trace. All information required to perform this simulation is present in the \ck{} file, with the
exception of a chemical kinetic model for \ce{H2}\slash \ce{CO} combustion.

In Python, additional functionality can be imported into a script or session by the \pybox|import|
keyword. Cantera and PyKED both must be imported into the session so that we can work with the code.
In the case of Cantera, we will use many functions from the library, so we assign it an abbreviation
(\pybox|ct|) for convenience. From PyKED, we will only be using the \pybox|ChemKED| class, so this
is all that is imported:
%
\begin{pythonbox}
import cantera as ct
from pyked import ChemKED
\end{pythonbox}
%
Next, we have to load the ChemKED file and retrieve the first element of the \pybox|datapoints|
list. Although this file only encodes a single experiment, the \pybox|datapoints| attribute will
always be a list (in this case, of length 1). As mentioned previously, the elements of the
\pybox|datapoints| list are instances of the \pybox|DataPoint| class, which we store in the variable
\pybox|dp|.
%
\begin{pythonbox}
ck = ChemKED('testfile_rcm.yaml')
dp = ck.datapoints[0]
\end{pythonbox}
%
The initial temperature, pressure, and mixture composition can be read from the
instance of the \pybox|DataPoint| class. \pk{} uses instances of the Pint \pybox|Quantity| class to
store values with units, while Cantera expects a floating-point value in SI
units as input. Therefore, we use the built-in capabilities of Pint to convert
the units from those specified in the \ck{} file to SI units, and we use the \pybox|magnitude|
attribute of the \pybox|Quantity| class to take only the numerical part. We also retrieve the
initial mixture mole fractions in a format Cantera will understand:
%
\begin{pythonbox}
T_initial = dp.temperature.to('K').magnitude
P_initial = dp.pressure.to('Pa').magnitude
X_initial = dp.get_cantera_mole_fraction()
\end{pythonbox}
%
With these properties defined, we have to create the objects in Cantera that represent the physical
state of the system to be studied.
